---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r echo=TRUE}
    activity <- read.table("data/activity.csv", sep=",", header=TRUE)
```


## What is mean total number of steps taken per day?

*Question 1 : total number of steps per day*
```{r echo=TRUE}
    subactivity <-  activity[complete.cases(activity),]
    stepbyday <- aggregate(subactivity$steps, by=list(subactivity$date), FUN=sum)
    colnames(stepbyday) <- c("date", "sumsteps")
    print(stepbyday)
```

*Question 2 : Histogram of total number each day*
```{r echo=TRUE}    
    hist(stepbyday$sumsteps, col = "green", main = "Total number of steps taken each day", xlab="Number of steps", breaks=20)
```

*Question 3 : Mean and Medium of the total number of steps per day*
```{r echo=TRUE}
    #aggregate(activity$steps, by=list(activity$date), FUN=function(x) c(mean, median),  na.rm=TRUE)
    mean(stepbyday$sumsteps, na.rm=TRUE)
    median(stepbyday$sumsteps, na.rm=TRUE)
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
